Inattentive responding can induce spurious associations between task behavior and symptom measures

### 2021-03-29 manuscripts/main.tex
checkpointing progress

% set up the author block
\setlength{\affilsep}{2em}   % set the space between author and affiliation
\author[1]{Samuel Zorowitz}
\author[1,2]{Daniel Bennett}
\author[1,3]{Yael Niv}
\affil[1]{Princeton Neuroscience Institute, Princeton University, USA}
\affil[2]{Department of Psychiatry, Monash University, Australia}
\affil[3]{Department of Psychology, Princeton University, USA}

% specify the title
\title{Inattentive responding can induce spurious associations between task behaviour and symptom measures}

@@

\begin{document}

@@

% make abstract
\abstract{Abstract}

@@

@@

With the advantages of online data collection also come specific drawbacks. Since participants recruited from online labour markets are typically completing experiments in their homes, they may be more likely to be distracted or multi-tasking during an experiment. Similarly, unsupervised participants may be more likely to use careless/insufficient effort (C/IE) response strategies with the intention to minimise expenditure of time and cognitive effort (e.g., responding randomly on self-report surveys or behavioural tasks). Among researchers using online labour markets, a common view is that poor-quality data resulting from C/IE responding can simply be treated as a source of unsystematic measurement error that can be overcome with increased sample sizes \cite{gillan2016taking, chandler2020participant}. Similarly, it is common practice in online behavioural research to mitigate poor-quality data using the same screening methods that are typically used in in-person data collection (e.g., excluding participants who perform at- or below-chance on behavioural tasks).

Here we wish to draw attention to an underappreciated feature of online psychiatric research using self-report symptom surveys. In such surveys, participants rate their endorsement of various psychiatric symptoms and, since most individuals in the general population tend to endorse no or few symptoms in many symptom domains, the resulting ground-truth symptom distributions tend to be heavily positively skewed. In this situation, the assumption that C/IE responding merely increases unsystematic measurement noise becomes untenable. Because of the positive skew in the ground-truth symptom distribution, participants who respond carelessly to the symptom survey are more likely to report higher levels of symptom endorsement relative to participants who complete the survey accurately \cite{chandler2020participant, ophir2020turker}. Consequently, unless C/IE survey responses are carefully identified and removed, a considerable proportion of putatively symptomatic individuals in an online sample may, in fact, be participants who have not engaged with the experiment with sufficient effort.

When participants complete both symptom surveys and behavioral tasks (a common study design in computational psychiatry), this artefact has the potential to induce spurious correlations between symptom self-reports and task behavior. Concretely, if the same participants who engage in C/IE responding on surveys (and who therefore inaccurately report high levels of psychiatric symptoms) also respond carelessly on behavioral tasks, then we would expect to observe an entirely spurious correlation emerge between greater symptom severity and worse task performance (see Figure \ref{fig:simulation}). A similar effect has been well documented in personality psychology, where the presence of careless respondents can induce correlations between questionnaires and bias estimated factors in factor analysis \cite{huang2012detecting, robinson2014inaccurate, huang2015insufficient, chandler2020participant, arias2020little}. To our knowledge, however, the potential for C/IE responding to induce spurious correlations between symptom endorsement and task behavior has not previously been noted in the burgeoning field of computational psychiatry.

\begin{figure}[t]
\includegraphics[width=16cm]{../figures/main_01.png}
\centering
\caption{Caption TK here}
\label{fig:simulation}
\end{figure}

Here we demonstrate the real risk that C/IE responding can lead to spurious symptom-task correlations in computational psychiatry research. First, we asked to what extent researchers in computational psychiatry screen self-report data from online samples. We found that the majority of these studies did not screen participants' survey data at all, and that very few followed best-practice recommendations for survey data screening. We then asked whether behavioural screening alone was sufficient to identify participants engaging in C/IE responding on psychiatric symptom surveys. In a new dataset from two separate online labour marketplaces, we found that screening based on task behaviour fails to adequately identify participants engaging in C/IE responding on surveys. Lastly, we investigated whether, under these circumstances, C/IE responding led to spurious correlations between symptom severity and task performance for positively-skewed symptom measures. Consistent with the logic set out above, we confirmed that failure to appropriately screen out C/IE survey responding in the proof-of-concept dataset that we collected would have produced a number of spurious correlations between task behaviour and self-reported symptoms of depression and anxiety.

@@

When the rate of symptom endorsement is low in the general population, participants who engage in C/IE responding on self-report symptom measures may appear as putative symptomatic individuals. If the same participants also perform a behavioural task with insufficient effort, they may in turn induce spurious correlations between measures of task behavior and self-report symptoms. The risk for erroneous inferences, however, is attenuated if experimenters identify and remove those participants. As such, it is imperative for researchers to screen for C/IE responding in their data. The extent to which experimenters evaluate the quality of both the task and self-report data they collect is unclear. We suspect that whereas screening of task data is standard practice, the screening of self-report data is far less prevalent in online human behavioral research.

To test the empirical support for this claim, we performed a narrative literature review of online human behavioral studies (see Methods for details of the literature search and the Github repository for the full list of included studies). Briefly, we identified studies for inclusion through searches on Google Scholar using permutations of query terms related to online labour platforms, experimental paradigms, and symptom measures. We included studies that (a) recruited participants online, (b) measured behavior on at least one experimental task, and (c) measured responses on at least one self-report symptom measure. We then evaluated whether and how each of the resulting 49 studies performed task and self-report data screening. We note that this review was not meant to be systematic, but instead to provide a representative overview of common screening methods in online behavioral studies.

Summary statistics concerning the prevalence and types of screening methods are provided in Table~\ref{tab:screening}. Of the 49 online behavioral studies meeting inclusion criteria, roughly 80\% (39/49) used at least one method to identify C/IE responding in task behavior. Of those studies, just over half relied on a single screening method, with considerable heterogeneity in behavior screening methods across studies. Most common (46\% of all studies) was identifying participants whose performance was indistinguishable from chance-level on some measure of accuracy. Almost as common (38\%) was screening based on low response variability (i.e., excluding participants who predominantly responded in the same fashion across trials, such as using only a single response key).

In contrast, only a minority (19/49, or 39\%) of online behavioral studies in our sample screened for C/IE responding in self-report symptom measures. Of those studies the most common method was the use of attention checks, which operate under the assumption that certain responses to these prompts are unlikely under attentive responding, such that participants who do not give the correct response are likely to be engaged in C/IE responding. Attention checks can be subdivided into instructed items (in which participants are explicitly told which response to select; e.g., `Please select ``Strongly Agree"'), and infrequency items (in which some responses are logically invalid; e.g., endorsing 'Agree' for the question `I competed in the 1917 Summer Olympic Games'). Of those studies that specified what type of attention check was used, instructed items were the most common method. As we discuss further below, this is notable because best-practice recommendations for data collection in personality psychology explicitly counsel against using instructed-item attention checks. Only a handful of studies employed statistical or so-called unobtrusive screening methods such as outlier detection or personal consistency.

In sum, screening methods for C/IE responding vary widely across the studies that we reviewed. Whereas screening for C/IE responding in task behavior is relatively common for online behavioral studies, screening of self-report survey data is far less prevalent. Though at first this pattern may seem troubling, low rates of survey data screening are not necessarily an issue if screening on task behavior alone is sufficient to remove participants engaging in C/IE responding. That is, screening on survey data may be redundant if there is a high degree of correspondence between task- and survey-based screening methods. To our knowledge, however, this assumption has never been empirically tested. In the next section, we explicitly test this hypothesis in a large sample of online participants completing a battery of self-report surveys and a behavioural task.

\begin{table}[t]
\centering
\setlength{\tabcolsep}{8pt}
\begin{tabular}{lllllll}
\toprule
& \multicolumn{2}{c}{Task Screening} & \multicolumn{4}{c}{Self-Report Screening} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-7}
Frequency & \multicolumn{2}{c}{N=39 (80\%)} & \multicolumn{4}{c}{N=19 (39\%)} \\
\midrule
Measure & Accuracy & 18 (37\%) & \multicolumn{2}{l}{Attention Check} & & 17 (35\%) \\
& Variability & 15 (31\%) & & \textit{Instructed} & & 10 (20\%) \\
& Response Time & \ 7 (14\%) & & \textit{Unspecified} & & 5 (10\%) \\
& Comprehension Check & \ \ 5 (10\%) & \multicolumn{2}{l}{Unobtrusive} & & \ 4 (8\%) \\
& Other & 16 (33\%) \\
\bottomrule
\end{tabular}
\caption{The prevalence and types of task and self-report data screening practices in a sample (N=49) of online behavioral studies.}
\label{tab:screening}
\end{table}

Before continuing, we wish to briefly highlight two additional results from our literature review. First, we found that the majority of studies used only one single method to detect C/IE responding. This is worrisome from the perspective of signal detection theory. As has been previously reported \cite{desimone2018dirty}, measures of C/IE responding vary in their sensitivities: some measures will identify only the most careless participants, whereas others will flag many more participants. As such, using only a single measure runs the risk of false positives (in the case of high sensitivity) and false negatives (in the case of low sensitivity). Similarly, screening methods vary in their sensitivity to different patterns of C/IE responding \cite{curran2016methods}. Using only one measure may thus only identify a subset of careless respondents. As has been discussed elsewhere, the use of multiple screening methods is encouraged to more accurately and sensitively identify C/IE responding in a dataset \cite{curran2016methods, barends2019noncompliant}.

Second, we found that the majority of studies that did screen self-report survey data relied on instructed attention checks. This is potentially problematic, since a growing body of research disputes the efficacy of instructed items as attention checks for online studies \cite{barends2019noncompliant, thomas2017validity, hauser2016attentive}. Indeed, instructed attention checks are often discussed on forums for workers of online labour markets\footnote{Several examples of such discussions can be found at the Github repository for this project:  \url{github.com/nivlab/silver-screen}}, and scripts have even been developed with the purpose to highlight words most commonly used in instructed attention checks. Unsurprisingly, then, it has been documented that subsets of participants may actively seek out instructed items in surveys and respond randomly elsewhere \cite{barends2019noncompliant}. In line with best-practice recommendations for online survey research, therefore, we encourage experimenters to use alternate means of screening for C/IE responding for self-report survey data (e.g., infrequency-item attention checks) \cite{huang2015detecting}.

@@

In the majority of the papers reviewed above, experimenters relied solely on behavioural screening to exclude participants who engaged in C/IE responding during an experiment. Nevertheless, even with behavioural screening, it is still possible that spurious behaviour-symptom correlations might emerge if there is low correspondence between behavioural and survey-based C/IE screening methods (since high correspondence between methods would entail that behavioural screening alone would exclude participants who would have failed survey screening).

In order to the measure the correspondence of screening measures estimated from task and self-report behavior, we conducted an online behavioral experiment involving a simple behavioral task and a battery of self-report psychiatric symptom measures (see Methods for details). A final sample of 386 participants from the Amazon Mechanical Turk (N=186) and Prolific (N=200) online labour markets completed a probabilistic reversal learning task and 5 self-report symptom measures. The reversal learning task required participants to learn through trial-and-error which of three choice options yielded reward most often, and was modeled after similar tasks used to probe reinforcement learning deficits in psychiatric disorders \cite{huang2017computational, waltz2007probabilistic, mukherjee_reward_2020}. The self-report measures assayed psychiatric symptoms related to depression, mania, anxiety, reward motivation, and anhedonia, and were chosen based on previous literature to have a variety of expected response distributions (symmetric and asymmetric). In line with current best-practice recommendations in personality psychology, each self-report instrument included one `infrequency' item that could be used to identify C/IE responses in survey data (see Methods for a list of infrequency items). To minimize any influence of fatigue, the entire experiment (surveys \& task) was designed to require 10 minutes on average to complete (observed mean = 10.28 minutes). For similar reasons, participants completed the surveys prior to beginning the task.

To assess the overall quality of the data, we examined the number of participants flagged by the choice accuracy and infrequency item screening measures (i.e. two most common screening measures identified in our literature review). Only 26 participants (7\%) were flagged as exhibiting choice behavior in the reversal learning task at or below chance levels. In contrast, 85 participants (22\%) responded suspiciously on one or more of the infrequency items when completing the self-report symptom measures. This discrepancy in the proportion of participants flagged by each method is consistent with previous research which have found varying levels of sensitivity to C/IE responding across screening methods \cite{desimone2018dirty}. Though we will not address platform differences here, we note the proportions of flagged participants were marginally but significantly greater on Mechanical Turk compared to Prolific for both accuracy (MTurk: N=18/186; Prolific: N=8/200; $z=2.22, p=0.026$) and infrequency (MTurk: 50/186; Prolific: 35/200; $z = 2.22, p = 0.026$) measures.

\begin{figure}[t!]
\includegraphics[width=16cm]{../figures/main_02a.png}
\centering
\caption{Caption TK here}
\label{fig:distributions}
\end{figure}

According to our hypothesis, a prerequisite for spurious behavior-symptom correlations to emerge is a mean-shift in the average level of symptom endorsement in participants engaging in C/IE responding relative to attentive participants, which might occur when the overall rate of symptom endorsement is low. As such, we visually inspected for each of the self-report measures the distribution of symptom scores in participants flagged by the infrequency measure vs. those that were not (hereafter referred to as `careless' or `attentive`, respectively; Figure \ref{fig:distributions}). In line with our predictions, the average level of symptom endorsement was noticeably exaggerated in careless participants relative to attentive participants for the symptom measures where symptom scores were positively-skewed (e.g. 7-up, 7-down, GAD-7). In contrast, where there was a greater range of symptom endorsement, the distributions of symptom scores between the two groups of participants were less noticeably distinct.

\begin{table}[!t]
\centering
\setlength{\tabcolsep}{6pt}
\begin{tabular}{lccccccc}
\toprule
& & \multicolumn{2}{c}{Total Score} & \multicolumn{2}{c}{Clinical Levels} & \multicolumn{2}{c}{Cronbach's $\alpha$} \\
\cmidrule(lr){3-4} \cmidrule(lr){5-6} \cmidrule(lr){7-8}
 Subscale & Skew & Attentive & Careless & Attentive & Careless & Attentive & Careless \\
\midrule
\midrule
7u    &  0.81 &    3.87 &   10.15 &     0.04 &     0.45 &    0.84 &    0.84 \\
7d    &  0.76 &    4.75 &   10.68 &     0.09 &     0.46 &    0.94 &    0.88 \\
GAD7  &  0.75 &    4.86 &    9.73 &     0.17 &     0.56 &    0.92 &    0.87 \\
BIS   &  0.78 &    7.74 &    7.92 &        - &        - &    0.83 &    0.62 \\
BAS   &  0.17 &   15.73 &   16.16 &        - &        - &    0.84 &    0.71 \\
SHAPS &  0.26 &    8.02 &   10.85 &     0.15 &     0.29 &    0.90 &    0.81 \\
PSWQ  &  0.19 &    4.78 &    6.74 &     0.07 &     0.08 &    0.93 &    0.81 \\
\bottomrule
\end{tabular}
\caption{}
\label{tab:screening}
\end{table}

To verify the infrequency item method as an accurate assay of C/IE responding, we performed three validation analyses. The first approach is to compare the proportion of participants meeting the cutoff for clinical psychopathology among careless versus attentive participants(Table \ref{tab:screening}). On the most positively-skewed measures, the fraction of participants scoring at or above clinical levels of symptom endorsement was dramatically inflated in careless responders. For example, previous studies of prevalence of generalized anxiety disorder found that roughly 5\% of the general population met the clinical threshold at a score of 10 or higher on the GAD-7 \cite{lowe2008validation, hinz2017psychometric}. Though clinical thresholds have not been established for the 7-up/7-down, recent work has suggested scores of 12 or higher as useful for discriminating between those with and without clinical levels of pathology \cite{youngstrom2020evaluating}. Though the fraction of participants meeting threshold for these measures in the attentive group was within the expected range of prevalence given epidemiological data, the exaggerated rates in the flagged group are suggestive of C/IE responding.\footnote{Interestingly, compared to previous literature the proportion of participants meeting threshold on the GAD-7 was elevated. We suspect this may reflect elevated rates of state anxiety during the COVID-19 pandemic, when this data was collected.}

The second approach was to compare the estimated internal consistency of the self-report measures between the careless and attentive groups. The logic is that, if C/IE responding manifests as a tendency to respond randomly, we should expect to see a decrease in the consistency of a measure in the careless group (as has been previously reported; \cite{maniaci2014caring, desimone2018dirty, desimone2018differential}). Accordingly, on the majority of measures we observed a reduction in Cronbach's $\alpha$ in the careless group relative to the attentive group (Table \ref{tab:screening}). A permutation test confirmed that the average decrease in internal consistency across measures was greater than would be expected by chance even controlling for the difference in the number of participants between the two groups ($t = 4.25, p = 0.006$).

The third and final approach is to statistically quantify the degree to which participants responded to the items in the self-report symptom inventory in a stereotyped fashion; that is, to determine if participants exhibited patterns in their responses that were independent of the contents of the items. To do so, we fit a random intercept item factor analysis model \cite{maydeu2006random} to the self report data (see Methods for details). Briefly, an intercept term is estimated for each participant which quantifies their bias towards using particular response options (e.g. leftmost, rightmost) regardless of what that option signifies for a particular self-report measure. We observed a credible difference between the average value of this intercept between the two groups (stats here), such that participants suspected of C/IE responding were biased towards using the right-half of the response options on the self-report symptom inventory. This translates to a tendency to endorse greater illness on the 7-up/7-down and GAD-7 (where the rightmost options indicate greater frequency of symptoms) but translates to a tendency to endorse \emph{less} illness on the SHAPS and BIS (where the rightmost options indicate lower frequency of symptoms) despite these inventories measuring strongly correlated constructs (i.e. depression and anhedonia, anxiety and behavioral inhibition). This pattern also helps explain why the distributions of symptom scores between attentive and careless participants on the SHAPS appear to be similar despite the low rates of symptom endorsement usually observed on the SHAPS \cite{snaith1995scale}.

>>>>

% set up the author block
\setlength{\affilsep}{2em}   % set the space between author and affiliation
\author[1]{Samuel Zorowitz}
\author[1,2]{Yael Niv}
\author[1,3]{Daniel Bennett}
\affil[1]{Princeton Neuroscience Institute, Princeton University, USA}
\affil[2]{Department of Psychology, Princeton University, USA}
\affil[3]{Department of Psychiatry, Monash University, Australia}

% specify the title
\title{Inattentive responding can induce spurious associations between task behavior and symptom measures}

@@

\begin{document}
%TC:ignore

@@

% make abstract
\abstract{
A common research design in the field of computational psychiatry involves leveraging the power of online participant recruitment to assess correlations between behaviour in cognitive tasks and the self-reported severity of psychiatric symptoms in large, diverse samples. Although large online samples have many advantages for psychiatric research, some potential pitfalls of this research design are not widely understood. Here we detail circumstances in which entirely spurious correlations may arise between task behaviour and symptom severity as a result of inadequate screening of careless or low-effort responding on psychiatric symptom surveys. Specifically, since many psychiatric symptom surveys have asymmetric ground-truth score distributions in the general population, participants who respond carelessly on these surveys will show apparently elevated symptom levels. If these participants are similarly careless in their task performance, and are not excluded from analysis, this may result in a spurious association between greater symptom scores and worse behavioural task performance. Here, we demonstrate exactly this pattern of results in $N = 386$ participants recruited online to complete a self-report symptom battery and a short reversal-learning choice task. We show that many behaviour-symptom correlations are entirely abolished when participants flagged for careless responding on surveys are excluded from analysis. We also show that exclusion based on task performance alone is not sufficient to prevent these spurious correlations. Of note, we demonstrate that false-positive rates for these spurious correlations \emph{increase} with sample size, contrary to common assumptions. We offer guidance on how researchers using this general experimental design can guard against this issue in future research; in particular, we recommend the adoption of screening methods for self-report measures that are currently uncommon in the field.
}

@@

%TC:endignore

@@

With the advantages of online data collection also come specific drawbacks. Since participants recruited from online labour markets are typically completing experiments in their homes, they may be more likely to be distracted or multi-tasking during an experiment. Similarly, unsupervised participants may be more likely to use heuristic response strategies with the intention to minimise expenditure of time and cognitive effort (e.g., responding randomly on self-report surveys or behavioral tasks). Here, we will refer to such inattentive or low-effort behaviors as careless/insufficient effort (C/IE) responding \cite{huang2012detecting, curran2016methods}. Among researchers using online labour markets, a common view is that poor-quality data resulting from C/IE responding can simply be treated as a source of unsystematic measurement error that can be overcome with increased sample sizes \cite{gillan2016taking, chandler2020participant}. Similarly, it is common practice in online behavioral research to mitigate poor-quality data using the same screening methods that are typically used in in-person data collection (e.g., excluding participants who perform at- or below-chance on behavioral tasks).

Here we wish to draw attention to an underappreciated feature of online psychiatric research using self-report symptom surveys. In such surveys, participants rate their endorsement of various psychiatric symptoms and, since most individuals in the general population tend to endorse no or few symptoms in many symptom domains, the resulting ground-truth symptom distributions tend to be heavily positively skewed \cite{lowe2008validation, tomitaka2018distributional}. In this situation, the assumption that C/IE responding merely increases unsystematic measurement noise becomes untenable. Because of the positive skew in the ground-truth symptom distribution, participants who respond carelessly to the symptom survey are more likely to report higher levels of symptom endorsement relative to participants who complete the survey accurately \cite{chandler2020participant, ophir2020turker}. Consequently, unless C/IE survey responses are carefully identified and removed, a considerable proportion of putatively symptomatic individuals in an online sample may, in fact, be participants who have not engaged with the experiment with sufficient attention or effort.

When participants complete both symptom surveys and behavioral tasks---a common study design in computational psychiatry---this artefact has the potential to induce spurious correlations between symptom self-reports and task behavior. That is, while C/IE behavior is traditionally thought of as a source of noise that can result in type II (false negative) errors, here we suggest that in large-scale online psychiatric studies it can instead result in type I (false positive) errors. Concretely, if the same participants who engage in C/IE responding on surveys (and who therefore inaccurately report high levels of psychiatric symptoms) also respond with insufficient effort on behavioral tasks, this can cause experimenters to observe an entirely spurious correlation between greater symptom severity and worse task performance (see Figure \ref{fig:simulation}). A similar effect has been well documented in personality psychology, where the presence of C/IE responding can induce correlations between questionnaires, and bias estimated factors in factor analysis \cite{huang2012detecting, robinson2014inaccurate, huang2015insufficient, chandler2020participant, arias2020little}.

\begin{figure}[t]
\includegraphics[width=16cm]{../figures/main_01.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{Simulated example of how spurious behavior-symptom correlations can arise when symptom endorsement is rare. \textit{Left:} When symptoms are moderately common in the general population, C/IE respondents (blue) are indistinguishable from attentive participants (red) in self-report measures (x-axis, marginal distribution shown on top). Despite the lower task performance of C/IE respondents (y-axis), no correlation arises between symptom scores and task performance (dots are participants drawn from the shown distributions, with \~15\% C/IE participants; dashed line shows the (lack of) correlation.) \textit{Right:} When symptoms are rare in the general population, careless respondents appear symptomatic in self-report measures. As a result, self-report symptom scores show a significant correlation with task performance.}
\label{fig:simulation}
\end{figure}

Here, we demonstrate the real risk that C/IE responding can lead to spurious symptom-task correlations in computational psychiatry research. First, we asked to what extent recent studies in computational psychiatry screen participants based on self-report symptom data. We found that the majority of these studies did not screen participants' survey data at all, and that very few followed best-practice recommendations for survey data screening \cite{barends2019noncompliant, thomas2017validity, hauser2016attentive}. We then asked whether behavioral screening alone was sufficient to identify participants engaging in C/IE responding on psychiatric symptom surveys. In a new dataset from two separate online labour marketplaces, we found that screening based on task behavior fails to adequately identify participants engaging in C/IE responding on surveys. Lastly, we investigated whether, under these circumstances, C/IE responding led to spurious correlations between symptom severity and task performance for positively-skewed symptom measures. Consistent with the logic set out above, we confirmed that failure to appropriately screen out C/IE survey responding in the proof-of-concept dataset that we collected would have produced a number of spurious correlations between task behavior and self-reported symptoms of depression and anxiety that are abolished when data are screened properly.

@@

First, we sought to what extent recent online studies screen participants in a way that would reduce the risk of spurious correlations due to C/IE participants. We performed a narrative literature review of 49 online human behavioral studies, and evaluated whether and how each study performed task and self-report data screening (see Methods for details of the literature search).

Among studies that we reviewed, approximately 80\% (39/49) used at least one method to identify C/IE responding in task behavior (see Table \ref{tab:screening}). Of these, just over half relied on a single screening method, with considerable heterogeneity in behavior screening methods across studies. Most common (46\% of all studies) was identifying participants whose performance was statistically indistinguishable from chance-level on some measure of accuracy. Almost as common (38\%) was screening based on low response variability (i.e., excluding participants who predominantly responded in the same fashion across trials, such as using only a single response key).

In contrast, only a minority (19/49, or 39\%) of studies screened for C/IE responding in self-report symptom measures. The most common survey screening method was the use of attention checks, which are prompts for which most responses are unlikely given attentive responding. Participants who do not give the correct response to these prompts are therefore likely to be engaged in C/IE responding. Attention checks can be subdivided into instructed items (in which participants are explicitly told which response to select; e.g., `Please select ``Strongly Agree"'), and infrequency items (in which some responses are logically invalid or exceedingly improbable; e.g., endorsing 'Agree' for the question `I competed in the 1917 Summer Olympic Games'). Of those studies that specified what type of attention check was used, instructed items were the most common method. As we discuss further below, this is notable because best-practice recommendations for data collection in personality psychology explicitly counsel \emph{against} using instructed-item attention checks \cite{barends2019noncompliant, thomas2017validity, hauser2016attentive}. Only a handful of studies employed statistical or so-called unobtrusive screening methods such as outlier detection or personal consistency.

In sum, whereas screening for C/IE responding in task behavior was relatively common for online behavioral studies, screening of self-report survey data was far less prevalent. Although this pattern may seem troubling, low rates of survey data screening are not necessarily an issue if screening on task behavior alone is sufficient to remove participants engaging in C/IE responding. That is, screening on survey data may be redundant if there is a high degree of correspondence between task- and survey-based screening methods.

In the next section, we explicitly test this hypothesis in a large sample of online participants completing a battery of self-report surveys and a behavioral task. Specifically, we measure the correspondence between the common task- and survey-based screening methods, as identified in our literature review, so that the results are informative with respect to the average online psychiatry study.

\begin{table}[t]
\centering
\small
\setlength{\tabcolsep}{8pt}
\begin{tabular}{lllrllr}
\toprule
& \multicolumn{2}{c}{Task Screening} & \multicolumn{4}{c}{Self-Report Screening} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-7}
Frequency & \multicolumn{2}{c}{N=39 (80\%)} & \multicolumn{4}{c}{N=19 (39\%)} \\
\midrule
Measure & Accuracy & 18 (37\%) & \multicolumn{2}{l}{Attention Check} & & 17 (35\%) \\
& Variability & 15 (31\%) & & Instructed & & 10 (20\%) \\
& Response Time & \ 7 (14\%) & & Infrequency & & 2 \, (4\%) \\
& Comprehension Check & \ \ 5 (10\%) & & Unspecified & & 5 (10\%) \\
& Other & 16 (33\%) & \multicolumn{2}{l}{Unobtrusive} & & \ 4 \, (8\%) \\
\bottomrule
\end{tabular}
\captionsetup{width=0.88\textwidth}
\caption{The prevalence and types of task and self-report data screening practices in a sample (N=49) of recent online behavioral studies.}
\label{tab:screening}
\end{table}

@@

%In the majority of the papers reviewed above, experimenters relied solely on behavioral screening to exclude participants who engaged in C/IE responding during an experiment. Nevertheless, even with behavioral screening, it is still possible that spurious behavior-symptom correlations might emerge if there is low correspondence between behavioral and survey-based C/IE screening methods (since high correspondence between methods would entail that behavioral screening alone would exclude participants who would have failed survey screening).

To measure the correspondence of screening measures estimated from task and self-report behavior, we conducted an online behavioral experiment involving a simple decision-making task and a battery of commonly used self-report psychiatric symptom measures (see Methods). A final sample of 386 participants from the Amazon Mechanical Turk (N=186) and Prolific (N=200) online labour markets completed a probabilistic reversal-learning task and 5 self-report symptom measures. The reversal-learning task required participants to learn through trial-and-error which of three options yielded reward most often, and was modeled after similar tasks used to probe reinforcement-learning deficits in psychiatric disorders \cite{waltz2007probabilistic, mukherjee_reward_2020}. The five self-report symptom measures were the 7-up (which measures symptoms of hypomania), the 7-down (which measures symptoms of depression), the GAD-7, (which measures generalized anxiety symptoms), the BIS/BAS (which measures reward and punishment motivations), the SHAPS (which measures anhedonia symptoms), and the PSWQ (which measures worry symptoms), and were chosen based on previous literature to have a variety of expected response distributions (symmetric and asymmetric). In line with current best-practice recommendations in personality psychology \cite{huang2015detecting}, each self-report instrument included one `infrequency' item that could be used to identify C/IE responses in survey data (see Methods for a list of infrequency items). The entire experiment (surveys and task) was designed to require 10 minutes on average to complete (observed mean = 10.28 minutes). To minimize any influence of fatigue on survey responding, participants completed the surveys prior to beginning the task.

%To assess the overall quality of the data, we examined the number of participants flagged by the choice accuracy and infrequency item screening measures (i.e. two most common screening measures identified in our literature review). 
Only 26 participants (7\%) were flagged as exhibiting choice behavior at or below statistically chance levels in the reversal-learning task. In contrast, 85 participants (22\%) endorsed a logically invalid or improbable response on one or more of the infrequency items when completing the self-report symptom measures. This discrepancy in the proportion of participants flagged by each method is consistent with previous research, which found varying levels of sensitivity to C/IE responding across screening methods \cite{desimone2018dirty}. The proportion of participants flagged for C/IE responding was marginally but significantly greater on Mechanical Turk compared to Prolific for both task (MTurk: N=18/186; Prolific: N=8/200; $z=2.22, p=0.026$) and survey data (MTurk: 50/186; Prolific: 35/200; $z = 2.22, p = 0.026$).

\begin{figure}[t!]
\includegraphics[width=16cm]{../figures/main_02a.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{Raincloud plots of total symptom scores in attentive (red) and C/IE (blue) participants. Each colored dot represents the symptom score for one participant. The black dots display the average score within each group (error bars denote 95\% bootstrap confidence interval). The scales are ordered according to their estimated skew from top-left (7-up) to bottom-right (PSWQ). The average level of symptom endorsement is most markedly different between groups in measures with lowest overall rates of endorsement.}
\label{fig:distributions}
\end{figure}

We hypothesise that spurious behavior-symptom correlations may emerge due to a mean-shift in the average level of symptom endorsement in participants engaging in C/IE responding relative to attentive participants. In turn, a mean-shift is expected to occur when the overall rate of symptom endorsement is low. To assess this, we visually inspected for each self-report measure the distribution of symptom scores in participants flagged by the infrequency measure vs. those that were not (Figure \ref{fig:distributions}). In line with our predictions, the average level of symptom endorsement was noticeably exaggerated in C/IE-responding participants for the symptom measures where symptom scores were most positively-skewed (7-up, 7-down, GAD-7; Table \ref{tab:validation}). In contrast, where there was higher rates of symptom endorsement overall, the distributions of symptom scores between the two groups of participants were less noticeably distinct.

\begin{table}[!t]
\centering
\small
\setlength{\tabcolsep}{6pt}
\begin{tabular}{lccccccc}
\toprule
& & \multicolumn{2}{c}{Total Score} & \multicolumn{2}{c}{Cronbach's $\alpha$} & \multicolumn{2}{c}{\% Clinical Cutoff} \\
\cmidrule(lr){3-4} \cmidrule(lr){5-6} \cmidrule(lr){7-8}
 Subscale & Skew & Attentive & C/IE & Attentive & C/IE & No Screening & With Screening \\
\midrule
7-up   &  0.81 &    3.87 &   10.15 &    0.84 &    0.84 &     13.0\% &      4.0\% \\
7-down &  0.76 &    4.75 &   10.68 &    0.94 &    0.88 &     17.4\% &      9.3\% \\
GAD-7  &  0.75 &    4.86 &    9.73 &    0.92 &    0.87 &     25.9\% &     17.3\% \\
BIS    &  0.78 &    7.74 &    7.92 &    0.83 &    0.62 &        - &        - \\
BAS    &  0.17 &   15.73 &   16.16 &    0.84 &    0.71 &        - &        - \\
SHAPS  &  0.26 &    8.02 &   10.85 &    0.90 &    0.81 &     17.9\% &     14.6\% \\
PSWQ   &  0.19 &    4.78 &    6.74 &    0.93 &    0.81 &      7.3\% &      7.0\% \\
\bottomrule
\end{tabular}
\captionsetup{width=0.88\textwidth}
\caption{Descriptive statistics of the self-report symptom measures between attentive and C/IE participants. Skew: the empirical skewness of the distribution of total symptom scores. Total score: the average symptom score between attentive and C/IE participants. Cronbach's $\alpha$: a measure of response consistency, where greater values closer to 1 indicate greater consistency in responses. \% Clinical Cutoff: the percentage of participants reaching threshold for clinical symptomology before and after screening based on the infrequency measure. (The BIS/BAS scales do not have clinical thresholds.)}
\label{tab:validation}
\end{table}

Hereafter, we use the infrequency-item method as a primary means of identifying C/IE responding in our data. To verify this approach, we conducted three validation analyses. The first analysis compared estimated internal consistency of self-report measures between the C/IE and attentive groups. The logic is that, if C/IE responding manifests as a tendency to respond randomly, we should expect to see a decrease in the consistency of a measure in the C/IE responding group \cite{maniaci2014caring, desimone2018dirty, desimone2018differential}. In line with this reasoning, we observed a reduction in Cronbach's $\alpha$ in the C/IE group for the majority of survey instruments (Table \ref{tab:validation}). A permutation test confirmed that the average decrease in internal consistency across measures was greater than would be expected by chance given the difference in participant numbers between groups ($t = 3.69, p = 0.021$).

Second, we quantified the degree to which participants responded to self-report symptom surveys in a stereotyped fashion; that is, we determined if participants exhibited patterns in their responses that were independent of the contents of the survey items. We fit a random-intercept item factor-analysis model \cite{maydeu2006random} to self-report data (see Methods), and for each participant we estimated an intercept parameter that quantified their bias towards using responses on the left or right side of the response scale, regardless of what that response signifies for a particular self-report measure (e.g., low on one symptom scale versus high on another). We observed a credible difference between the average value of this intercept for the two groups ($\Delta \text{intercept} = 0.67, 95\% \ \text{HDI} = [-0.76, -0.57]$), such that C/IE participants were biased towards using the right-half of survey response options. This translates to a tendency to endorse \emph{more severe} symptoms on the 7-up/7-down and GAD-7 scales (where the rightmost options indicate greater frequency of symptoms) but \emph{less severe} symptoms on the SHAPS and BIS (where the rightmost options indicate lower frequency of symptoms) despite these inventories measuring strongly correlated constructs (i.e. depression and anhedonia, anxiety and behavioral inhibition).
